---
title: "Laplacian matrix of a graph"
author: "Blake Baird"
date: "2/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem
Let A be the incidence matrix of a directed graph with n nodes and m edges. 
The **Laplacian matrix** associated with the graph is defined as $L = AA^T$ 
which is the Gram matrix of $A^T$. 

(a) Show that $D(v) = v^TLv$, where $D(v)$ is the Dirichlet energy defined on page 135.
(b) Describe the entries of $L$. 

Hint. The following two quantities might be useful: 
    - the degree of a node, which is the number of edges that connect to the node (in either
direction)
    - the number of edges that connect a pair of distinct nodes (in either
direction)

```{r chnk1}
library(igraph)
library(ggplot2)
g = igraph::graph_from_literal( A-+B, B-+D, B-+C, D-+A, D-+C)
plot(g,edge.label = c("1","2","3","4","5"))
```

The incidence matrix of this directed graph has n = 4 nodes and m = 5 edges.

```{r chnk2}
inc_mat = cbind(c(-1,1,0,0),c(0,1,0,-1),c(0,-1,1,0), c(1,0,0,-1), c(0,0,1,-1))
rownames(inc_mat) = c("A","B","C","D")
inc_mat
```
**Describe the entries of the Laplacian matrix**. The *degree of the node*, i.e. the number
of edges that connect to it, is shown on the diagonal. The off-diagonal entries 
are the negative of the number of edges connecting that distinct pair of nodes.

$L_{ii}|{i \in 1..n} = degree\_of\_node\_i$

$L_{ij}|{i \neq j} = -\ distinct\_edges\_beween\_i\_and\_j$
```{r chnk3}
L = inc_mat %*% t(inc_mat) 
L
```