---
title: "Integral of product of polynomials - Prob 10.26"
author: "Blake Baird"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Statement
Given two quadratic polynomials $p(x) = c_1 + c_2x + c_3x^2$ and 
$q(x) = d_1 + d_2x + d_3x^2$.
Express the integral $$J = \int_0^{1} p(x)q(x)dx$$ in the form $$J = c^{T}Gd$$

$c$ and $d$ are the coefficient vectors of p and q. Give the entries of G (a 3x3 matrix) as numbers.

## Solution

Find the value of the integral.
```{r chunk2}
library(Ryacas)
# Construct character matrix
p_x = "c1 + c2*x + c3*x^2"
q_x = "d1 + d2*x + d3*x^2"
p_x = ysym(p_x)
q_x = ysym(q_x)
prod = (p_x * q_x) %>% y_fn("Expand")
value_of_integral = prod %>% y_fn("Integrate(x,0,1)")
value_of_integral %>% y_fn("Simplify")
value_of_integral %>% y_fn("Factor")
```
If $J = c^T(Gd)$, we can see what the rows of $G' = (Gd)$ must be from the coefficients of 
$c_1$, $c_2$, and $c_3$ in the above, since $c^TG'$ is an inner product of 3-vectors. 
$$Gd = \begin{bmatrix}
d_1 + \frac{d_2}{2} + \frac{d_3}{3}\\
\frac{d_1}{2} + \frac{d_2}{3} + \frac{d_3}{4}\\
\frac{d_1}{3} + \frac{d_2}{4} + \frac{d_3}{5}
\end{bmatrix}$$

Applying G to the 3-vector d yields the above 3-vector, so G must be:
$$G = \begin{bmatrix}
1  & \frac{1}{2} & \frac{1}{3}\\
\frac{1}{2} & \frac{1}{3} & \frac{1}{4}\\
\frac{1}{3} & \frac{1}{4} & \frac{1}{5}
\end{bmatrix}$$

