---
title: "10.39 Gram Matrix and QR Factorization"
author: "Blake Baird"
date: '2022-03-28'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## QR Factorization Explanation

A is n x k with lin. indep. columns $a_1, a_2 ...$  
Q is n x k with columns $q_1, q_2, ...$ which form an orthonormal basis.  
$$Q^TQ = I$$


```{r c1}
library(pracma)
norm = function(X) { sqrt(sum(X * X)) }
A = matrix(c(2,-1,1,0),nrow=2,byrow=TRUE)
Q = gramSchmidt(A)$Q
R = gramSchmidt(A)$R
A
Q
R
```
Q comes from Gram-Schmidt-like process
```{r c2}
# q1 = normalized A1
q1tilde = A[,1]
q1 = q1tilde / norm(q1tilde)
q1

# q2tilde = a2 - projection of a2 on q1
q2tilde = A[,2] - c(crossprod(q1,A[,2]))*q1  
q2 = q2tilde/norm(q2tilde)  
q2
```


## Prob 10.39 What is relationship between gram matrix of A and the gram matrix of R?

$R = Q^TA$  
$R^TR = (Q^TA)^T(Q^TA)$  
$R^TR = A ( Q Q^T ) A^T$  

Thus, $$R^TR == A^TA$$.   

```{r c3}
( t(Q) %*% A ) - R
crossprod(A)
crossprod(R)
```

## What can you say about the angles between the columns of A and the angles between the columns of R?

Formula for the angle between two vectors is:
$\theta = acos(\frac{a_1^Ta_2}{||a_1||||a_2||})$  

Since the gram matrix of A and R are the same, then the norms and inner
products of their columns with their other columns are equal. i.e.,
$||a_1|| == ||r_1||\ \ and\ \ ||a_2|| == ||r_2||$  
and $a_1^T a_1 == r_1^T r_1$.  
  
Thus the angles between the columns of A and the angles between
the columns of R are the same.  
$$\angle(a_i , a_j) == \angle(r_i , r_j)$$




